Limits to Spontaneous Resets in the Double-Drift Illusion

### 2019-12-17 DoubleDriftResetLimits.Rmd
rewrote stats part for first bounded tracking experiment

# The functions in here are used on the data of the first tracking experiment:
source('R/tracking_bounded.R')
# This code is for the second tracking experiment:
source('R/tracking_reset_pause.R')
# And finally some code for the re-tracin experiment...
source('R/retrace_onePass_V4.R')
# As well as models fit to that data:
source('R/models.R')
```

@@

```{r}
# from R/data.R
downloadRawData(overwrite=FALSE, experiments=c('bounded_tracking', 'reset_pause', 'onePass_V4'))
```

@@

```{r fig.width=6, fig.height=9}
# from R/tracking_bounded.R:
plotBoundedTrackingRaw()
```

We can see that the orientation of the tracking movements responds to the strength and direction of the internal motion of the stimulus. However, there are no clear spontaneous resets of the stimulus.

# Experiment 1b: continuous tracking with pauses

Therefore, we did second experiment that mimicked the first, but with three changes, all designed to increase the occurrence of resets. First, the external motion was slowed down as earlier work showed that resets often occur after around 2 seconds, which was how long one pass (from one end of the screen to the other) took in the first experiment. Second, we only used -3 and 3 cycles per second for the internal motion. Third, on 3 of every 4 trials, a ~266.66 ms "blink" (8 frames at 30 Hz) was introduced in the stimulus. During this period the stimulus disappeared from the screen and reappeared after the blink in the same location to continue as before. The blinks occured equally often in the first, second or third full pass of the stimulus, and always halfway along the path on screen. These blinks have been shown to trigger resets [@Nakayama2018], so if we can spot triggered resets in the tracking data, we can use the characteristics of that tracking data to detect spontaneous resets. However, we would primarily expect to see increases in horizontal tracking speed, perhaps accompanied by a decrease in vertical speed, when people track the "jump" in the perceived location of the stimulus.

>>>>

# The functions in here are used on the data of the bounded tracking experiments:
source('R/tracking_bounded.R')
# And finally some code for the re-tracing experiment...
source('R/retrace_onePass_V4.R')
# As well as models fit to that data:
source('R/models.R')
```

@@

```{r}
# from R/data.R
downloadRawData(overwrite=FALSE, experiments=c('bounded_tracking', 'bounded_tracking_V2', 'onePass_V4'))
```

@@

```{r fig.width=7, fig.height=6}
# from R/tracking_bounded.R:
plotBoundedTracking()
```

The top row shows average instantaneous heading (as actual direction) over time (distance from origin), with a polar heatmap of all the data in the background. The middle 1 second of each pass of the gabor is used, so that the inversions at the far and close end of the workspace are excluded from the average heading data. Since headin doesn't change throughout these segments (stats?) we take the average heading across all segments for every participant and every internal speed to do statistical analyses.

```{r}
analyzeBoundedTracking()
```

First we want to see if internal motion affects illusion strength as measured by instantaneous heading direction. A repeated measures ANOVA on a model using internal speed to predict heading direction shows that heading direction in tracking data is a good measure of illusion strength (F(4,12)=102.7, p<.001, ges=0.97, Greenhouse-Geisser corrected). This effect is not caused by one internal speed giving different directions, but by each internal speed resulting in different heading directions from it's direct neighbour (4 FDR-corrected, pairwise, one-sided t-tests: all p<.025).

A few of the tracking paths might be reflecting a reset (four of them are plotted in red in Fig 2). However, there are only a few and it's not clear they represent resets, or tracking errors made when changing direction at the far and close end of the workspace. The heading distributions don't seem to show any systematic deflections from the average heading, for eample evoked after a specific amount time.

# Experiment 1b: continuous tracking

Therefore, we did another experiment that mimicked the first, but with two changes, all designed to increase the occurrence of resets. First, we only used -3 and 3 cycles per second for the internal motion as this would increase the strength of the illusion and perhaps lead to a higher chance of reset. Second, the external motion was halved as earlier work showed that resets often occur after around 2 seconds [@Lisi2015 ??], and 2 seconds was how long one pass (from one end of the screen to the other) took in the first experiment. So now, the double-drift stimulus would do a single pass in 4 seconds instead of 2 seconds. This allows more opportunity for resets, but might also make it easier to track the gabor's position, so that errors at the direction changes in the far and close end of the workspace are less likely.
